---
title: "NEWS-MassIBItools"
output: github_document
---

<!-- NEWS.md is generated from NEWS.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "NEWS-"
)
```

```{r, echo = FALSE}
cat(paste0("Last Update: ",Sys.time()))
```

MassIBItools v0.1.2.920 (2020-12-07)
================
* Updated data prep, instructions, and background information per Jen Stamp's (Tt) edits

MassIBItools v0.1.2.915 (2020-11-02)
================
* Added HTML versions of instructions and background tabs


MassIBItools v0.1.2.900 (2020-09-24)
================
* Added instructions tab that has background information

MassIBItools v0.1.1.900 (2020-09-23)
================
* Added data explorer tab - includes increased mapping capabilities and index scoring plot. 
* Added generic summary report that downloads when results are downloaded.

MassIBItools v0.1.0.904 (2020-09-11)
================
* Created additions to mappping capabilities - site selector, toggle of background layers, and region layers. 

MassIBItools v0.1.0.900 (2020-09-08)
================

* Forked code from https://github.com/Blocktt/MassIBItools
